\LARGE \bf Imposing a Weight Norm Constraint for Neuro–Adaptive Control

### 2025-03-11 manuscript.tex
bold vec. and mat.

\let\proof\relax
\let\endproof\relax
\usepackage{amsthm}
\newtheorem{remark}{Remark}
\newtheorem{assum}{Assumption}
\newtheorem{lem}{Lemma}
\newtheorem{theorem}{Theorem}

\makeatletter
\let\NAT@parse\undefined
\makeatother
\usepackage{hyperref}
\hypersetup{
    linkbordercolor=blue,
    filebordercolor=red,
    citebordercolor = green,      
    urlbordercolor=red,
    %
    % colorlinks=true,
    % linkcolor=blue,
    % filecolor=blue,
    % citecolor = blue,      
    % urlcolor=cyan,
}

@@

% One of the common issues in neuro–adaptive control is that the boundedness of the control inputs is not guaranteed.
One of the common issues in using NNs is that the outputs of NNs are not predictable.
This is because the input‒output relationships of the NNs are not interpretable (i.e., the NNs are usually called black boxes \cite{RN8, RN9}).
This issue should be addressed for safety because the unpredictable control input may reach the actuator's physical limitations and result in system collapse.

@@

\begin{itemize}
    \item $\otimes$ denotes the Kronecker product \cite{RN22}.
    \item $x_{(i)}$ denotes the $i\textsuperscript{th}$ element of vector $x$.    
    % \item $A_{(i,j)}$ is the $i\textsuperscript{th}$ row-$j\textsuperscript{th}$ column element of matrix $A$.
    \item $\text{row}_j(A)$ denotes the $j\textsuperscript{th}$ row of $A\in\mathbb{R}^{n\times m}$. 
    % \item $\text{col}_i(A)$ and $\text{row}_j(A)$ denote the $i\textsuperscript{th}$ column and $j\textsuperscript{th}$ row of $A\in\mathbb{R}^{n\times m}$, respectively.
    \item $\text{vec}(A)\triangleq [\text{row}_1(A^T)  ,\cdots,\text{row}_m(A^T)  ]^T   $ for $A\in\mathbb{R}^{n\times m}$.
    % \item $\text{vec}(A)\triangleq [\text{col}_1(A)^T  ,\cdots,\text{col}_m(A)^T  ]^T   $ for $A\in\mathbb{R}^{n\times m}$.
    \item $\lambda_\text{min}(A)$ denotes the minimum eigenvalue of $A\in\mathbb{R}^{n\times n}$.
\end{itemize}

@@

Consider an uncertain Euler‒Lagrange system modeled as
\begin{equation}
    M(q)\ddot q + C(q,\dot q)\dot q + G(q) + F(q) = \tau
    \label{eq. system dynamics 1}
\end{equation}
where $q\in \mathbb{R}^n$ and $\tau\in\mathbb R^n$ denote the generalized coordinate and the control input, respectively; $M(q)\in\mathbb{R}^{n\times n}$, $C(q,\dot q)\in\mathbb{R}^{n\times n}$, and $G(q)\in\mathbb{R}^{n}$ denote the unknown system function matrices; and $F(q)\in\mathbb{R}^{n}$ denotes the external force.
Using the user-designed matrices $M_0>0,C_0$ and $G_0$, \eqref{eq. system dynamics 1} can be represented as 
\begin{equation}
    M_0\ddot q+C_0\dot q+G_0 = \tau + f(q,\dot q,\ddot q)
    \label{eq. system dynamics 2}
\end{equation}
where $f(q,\dot q,\ddot q) \triangleq -(M-M_0)\ddot q-(C-C_0)(q,\dot q)\dot q -(G-G_0) -F(q)$ denotes the residual unknown term.

Hence, the objective of the control design is to make $q$ track the continuously differentiable desired trajectory $q_d(t):\mathbb R\to \mathbb R^n$ under the unknown terms $f$.

@@

The backstepping control-based approach is utilized to generate a reference signal $z^*\triangleq -{k_q}\tilde q+\dot q_d$ for ${z}\triangleq \dot q$, where $\tilde q\triangleq {q} - q_d$ and $k_q\in\mathbb R_{>0}$.
The desired stabilizing controller can be designed as follows:
\begin{equation}
    \tau^* = -M_0\cdot ({k_z}\tilde z)+ 
        ( 
            -M_0\tilde q+C_0{z}+G_0-f+M_0 \dot z^*
        )
    \label{eq. desired control}
\end{equation}
where $\tilde z\triangleq z-z^*$ and $k_z\in\mathbb R_{>0}$.
Note that the desired controller cannot be realized because $f$.

To approximate the desired controller, an NN is utilized.
The NN with a single hidden layer is represented as 
\begin{equation*}
    \Phi({q_{NN}};\theta)\triangleq V_1^T\phi(V_0^T{q_{NN}})
\end{equation*}
where ${q_{NN}}\in\mathbb R^{l_0+1}$ denotes the NN input vector, $V_i\in\mathbb R^{(l_i+1)\times l_{i+1}},\ i\in[0,1]$ denotes the weight matrix of the $i\textsuperscript{th}$ layer and $\phi:\mathbb R^{l_1}\to\mathbb R^{l_1+1}$ denotes the activation function layer.
The activation function layer consists of an elementwise nonlinear function $\sigma(\cdot)$ and an augmented $1$ to combine the bias terms in the weight matrix (i.e., $\phi(x) = [\sigma(x_{(1)}),\cdots, \sigma(x_{(l_1)}), 1]^T$).
For further simplicity, let $\theta\triangleq[\theta_1^T,\theta_0^T]^T\in\mathbb R^{\Xi}$ denote the total weight vector, where $\theta_i\triangleq \text{vec}(V_i)\in\mathbb R^{\Xi_i}$ denotes the vectorized weights.
$\Xi_i=(l_i+1)\cdot l_{i+1}$ and $\Xi=\Xi_0+\Xi_1$ denote the number of each layer and total weights, respectively.

Using this NN, the desired controller $\tau^*$ can be approximated by the ideal weight vector $\theta^*$ for a compact subset $\Omega_{NN}\in\mathbb R^{l_0+1}$ to $\epsilon$ accuracy such that $\sup_{{q_{NN}}\in\Omega_{NN}} \Vert\Phi({q_{NN}};\theta^*) - \tau^*\Vert=\epsilon<\infty$ \cite{RN35}.
% The boundedness of the norm of $\theta^*$  and approximation error $\epsilon$ are also claimed in \cite{RN35}.
The ideal weight vector $\theta^*$ is typically assumed to be bounded.
Then, using the estimated weight vector $\hat\theta=[\hat\theta_1^T,\hat\theta_0^T]^T$ of $\theta^*=[\theta_1^{*T},\theta_0^{*T}]^T$, the desired controller $\tau^*\approx -\Phi(q_{NN};\theta^*)-\epsilon$ can be approximated as follows:
\begin{equation}
    \tau \triangleq -\Phi({q_{NN}};\hat\theta).
    \label{eq. control law}
\end{equation}
For further sections, let $\Phi^*\triangleq\Phi({q_{NN}};\theta^*)$ and $\phi^* \triangleq\phi(V_0^{*T}{q_{NN}})$, and let $\hat\Phi\triangleq\Phi({q_{NN}};\hat\theta)$, $\hat\phi \triangleq\phi(\hat V_0^{T}{q_{NN}})$ and $\hat\phi' \triangleq \partial \hat\phi/\partial (\hat V_0^Tq_{NN})$.

Using \eqref{eq. system dynamics 2}, \eqref{eq. desired control}, and \eqref{eq. control law}, the error dynamics can be obtained as the first-order system of augmented error $\xi\triangleq[\tilde q^T,\tilde z^T]^T\in\mathbb R^{2n}$ as follows:
\begin{equation*}
    \dot\xi = A_\xi\xi+B_\xi(\Phi^*-\hat\Phi+\epsilon)
\end{equation*}
where 
\begin{equation*}
    A_\xi \triangleq 
    \begin{bmatrix}
        -{k_q} I_n &I_n\\-I_n& -{k_z} I_n
    \end{bmatrix}
    ,\ 
    B_\xi \triangleq 
    \begin{bmatrix}
        0_{n\times n}\\M_0^{-1}
    \end{bmatrix}.
\end{equation*}

@@

Using the augmented error $\xi$, the quadratic objective function is defined as
\begin{equation*}
    J(\xi;\hat\theta)\triangleq
    {1\over 2}
    \xi^TW\xi
\end{equation*}
where $W=W^T>0$ denotes the weighting matrix.

For the boundedness of the weights, weight norm constraints are imposed on the adaptation process such that $c_{\theta_i}=\Vert\hat\theta_i\Vert^2-\bar\theta^2_i,\ i\in[0,1]$.
% Then the objective of the adaptation process can be reformulated into the constrained optimization problem as follows:
The control problem can be reformulated into a constrained optimization problem as follows:
\begin{equation*}
    \begin{matrix}
        \min_{\hat\theta} \ J(\xi;\hat\theta)
        \\ \\
        \begin{aligned}
        \text{subject to\;}&c_{j}\le0, \quad j\in\mathcal A = [\theta_0,\theta_1].
        \end{aligned}
    \end{matrix}
\end{equation*}
In this optimization problem, $\xi$ is considered a predefined parameter.
The corresponding Lagrangian function is defined as
\begin{equation}
    L(\xi,\hat\theta,[\lambda_j]_{j\in\mathcal A})\triangleq J(\xi;\hat\theta)+\sum_{j\in\mathcal A} \lambda_jc_j(\hat\theta)
    .
\end{equation}
% where $\mathcal A=\{j\ \vert \ c_j\ge 0\}$ denotes the active set.
To solve the dual problem $\min_{\hat\theta} \max_{[\lambda_j]_{j\in\mathcal A}}  L$, the adaptation law is derived as follows:
\begin{align}
        \dot {\hat\theta}&=-\alpha {\partial L\over\partial \hat\theta}
        =-\alpha 
        \bigg(
        {\partial J\over \partial \hat\theta}+\sum_{j\in\mathcal{A}}
        \lambda_j {\partial c_j\over \partial \hat\theta}
        \bigg),
    \label{eq. adaptation law th}
        \\
        \dot\lambda_j& = \beta_j{\partial L\over\partial \lambda_j} = \beta_j c_j ,
        \quad\quad\quad\quad      \      
        \forall j\in\mathcal A,
    \label{eq. adaptation law L}
        \\
        \lambda_j & = \max(\lambda_j,0) ,
        \quad\quad\quad\quad\ \ \ \ \ 
        \forall j\in\mathcal A,
    \label{eq. adaptation law L max}
\end{align}
where $\alpha\in\mathbb R_{>0}$ and $\beta_j\in\mathbb R_{>0}$ denote the adaptation gain and the update rate for each Lagrange multiplier, respectively.
\color{red}
% The Eq.~\eqref{eq. adaptation law L max} is analog to using an active set $\mathcal A=\{j\ \vert \ c_j\ge 0\}$ in constrained optimization theory which excludes terms of satisfied constraints in adaptation law.
\color{black}

Using the chain rule, the gradient of the objective function with respect to the weights (i.e., $\partial J/\partial \hat\theta$) in \eqref{eq. adaptation law th} can be represented as
\begin{equation*}
    {\partial J\over\partial \hat\theta}
    =
    {\partial \xi\over\partial \hat\theta}^TW
    {\partial \xi\over\partial \hat\theta}
    .
\end{equation*}
The calculation of $\partial J/\partial \hat\theta$ is not straightforward because of the dynamics of $\xi$.
Using the forward sensitivity method presented in \cite{RN34}, the sensitivity equation can be obtained as follows:
\begin{equation}
    \dot \eta =A_\xi - B_\xi 
    % (\partial \hat\Phi/\partial \hat\theta)
    {\partial \hat\Phi\over\partial \hat\theta}
    \label{eq. sen eq}
\end{equation}
where $\eta\triangleq \partial \xi/\partial \hat\theta\in\mathbb R^{2n\times \Xi}$ denotes the sensitivity of the weights to the augmented error.
The initial value of $\eta$ is zero since the initial $\xi$ is independent of the weights.
By decomposing for each layer, the dynamics of $\eta_i\triangleq \partial \xi/\partial \hat\theta_i\in\mathbb R^{2n\times \Xi_i}$ can be represented as
\begin{equation*}
    \begin{aligned}
        \dot\eta =&
        \begin{bmatrix}
            \eta_1&\eta_0
        \end{bmatrix}'
        \\
        =&A_\xi
        \begin{bmatrix}
            \eta_1&\eta_0
        \end{bmatrix}
        -B_\xi
        \begin{bmatrix}
            (I_{l_2}\otimes \hat\phi^T)&\hat V_1^T\hat\phi'(I_{l_1}\otimes {q_{NN}}^T)
        \end{bmatrix}
        .
    \end{aligned}
\end{equation*}
The calculation of $\partial \hat\Phi/\partial\hat\theta$ is introduced in \cite{RN36}.
In conclusion, the gradient of the objective with respect to the weights can be obtained as ${\partial J/\partial \hat\theta} = \eta^TW\xi$ by simulating the sensitivity equation \eqref{eq. sen eq}.

On the other hand, the gradient of the constraints with respect to the weights can be represented as follows:
\begin{equation*}
    \frac{\partial c_{\theta_0}}{\partial\hat\theta}
    = 
    \begin{bmatrix}
        0_{\Xi_1\times 1} \\
        2\hat\theta_0 
    \end{bmatrix}
    ,\quad
    \frac{\partial c_{\theta_1}}{\partial\hat\theta}    
    = 
    \begin{bmatrix}
        2\hat\theta_1 \\
        0_{\Xi_0\times 1}
    \end{bmatrix}
    .
\end{equation*}

@@

\begin{theorem}
For the dynamical system in \eqref{eq. system dynamics 1}, the proposed controller \eqref{eq. control law} and the adaptation law \eqref{eq. adaptation law th}, \eqref{eq. adaptation law L} and \eqref{eq. adaptation law L max} ensure the boundedness of the tracking error $\xi$ and the weight estimation $\hat\theta$, provided that the control gains ${k_q}$ and ${k_z}$ satisfy \eqref{eq. ctrl stable condition}.
\end{theorem}

@@

\subsection*{Step 1: Boundedness of $\hat\theta_1,\eta_1,\xi$}

@@

The dynamics of $\xi$ can be represented as
\begin{equation*}    
    \dot \xi = A_\xi\xi +B_\xi
    (
        -\hat V_1^T \hat\phi+w(t)
    )
\end{equation*}
where $w(t)\triangleq V_1^{*T}\phi^*+\epsilon$ is a lumped residual term, which is bounded as $\Vert w(t)\Vert\le \bar w< 0$ since $\Vert\theta_1^*\Vert$ and $\Vert\phi^*\Vert$ are bounded.
The dynamics of $\eta_1$ and $\hat\theta_1$ are represented as
\begin{equation*}
    \begin{aligned}
        \dot\eta_1 =
        & 
        A_\xi \eta_1 -B_\xi (I_{l_{2}}\otimes \hat\phi^T)
        \\
        \dot{\hat\theta}_1 =
        & -\alpha 
        (
            \eta_1^TW\xi+2\lambda_{\theta_1} \hat\theta_1
        ).
    \end{aligned} 
\end{equation*}
According to \cite[Chap.~4 T.~1.9]{RN32}, the boundedness of $\eta_1$ can be obtained since $A_\xi$ is stable and the residual term $-B_\xi (I_{l_{2}}\otimes \hat\phi^T)$ is bounded.

>>>>

\let\proof\relax
\let\endproof\relax
\usepackage{amsthm}

\makeatletter
\let\NAT@parse\undefined
\makeatother
\usepackage[colorlinks=true,linkcolor=black,citecolor=black]{hyperref}
% \usepackage[]{hyperref}
% \hypersetup{
%     % bookmarksnumbered=true,
%     %   linkbordercolor=blue,
%     %   filebordercolor=red,
%     %   citebordercolor = green,      
%     %   urlbordercolor=red,
%       %
%       colorlinks=false
%       % linkcolor=blue,
%       % filecolor=blue,
%       % citecolor = blue,      
%       % urlcolor=cyan,
% }

\usepackage{cleveref, autonum}

\newtheorem{remark}{Remark}
\newtheorem{assum}{Assumption}
\newtheorem{lem}{Lemma}
\newtheorem{theorem}{Theorem}

\newcommand{\etal}{\textit{et~al.\ }}
\newcommand\eg{\textit{e.g.\ }}
\newcommand\ie{\textrm{i.e.\ }}
\newcommand\R{\mathbb{R}}

\newcommand*{\mv}[1]{\boldsymbol{#1}}
\newcommand*{\mm}[1]{\boldsymbol{#1}}
\newcommand*{\norm}[1]{\lVert #1 \rVert}

\newcommand\bx{\mv x}
\newcommand\bu{\mv u}

\newcommand\der{\mathrm d}
\newcommand*{\ddfrac}[2]{
    \frac{\der {#1}}{\der {#2}}
}
\newcommand*{\ddtfrac}[2]{
    \tfrac{\der {#1}}{\der {#2}}
}

@@

% One of the common issues in neuro–adaptive control is that the boundedness of the control inputs is not guaranteed.
One of the common issues in using NNs is that the outputs of NNs are not predictable.
This is because the input‒output relationships of the NNs are not interpretable (\ie the NNs are usually called black boxes \cite{RN8, RN9}).
This issue should be addressed for safety because the unpredictable control input may reach the actuator's physical limitations and result in system collapse.

@@

\begin{itemize}
    \item $\otimes$ denotes the Kronecker product \cite{RN22}.
    \item $\mv x_{(i)}$ denotes the $i\textsuperscript{th}$ element of vector $\mv x$.    
    % \item $A_{(i,j)}$ is the $i\textsuperscript{th}$ row-$j\textsuperscript{th}$ column element of matrix $A$.
    \item $\text{row}_j(\mm A)$ denotes the $j\textsuperscript{th}$ row of $\mm A\in\mathbb{R}^{n\times m}$. 
    % \item $\text{col}_i(A)$ and $\text{row}_j(A)$ denote the $i\textsuperscript{th}$ column and $j\textsuperscript{th}$ row of $A\in\mathbb{R}^{n\times m}$, respectively.
    \item $\text{vec}(\mm A)\triangleq [\text{row}_1(\mm A^\top )  ,\cdots,\text{row}_m(\mm A^\top )]^\top $ for $\mm A\in\mathbb{R}^{n\times m}$.
    % \item $\text{vec}(A)\triangleq [\text{col}_1(A)^\top  ,\cdots,\text{col}_m(A)^\top  ]^\top   $ for $A\in\mathbb{R}^{n\times m}$.
    \item $\lambda_\text{min}(A)$ denotes the minimum eigenvalue of $A\in\mathbb{R}^{n\times n}$.
\end{itemize}

@@

Consider an uncertain Euler‒Lagrange system modeled as
\begin{equation}
    \mm M(\mv q)\ddot {\mv q} + \mm C(\mv q,\dot {\mv q})\dot {\mv q} + \mm G(\mv q) + \mm F(\mv q) = \mv \tau
    \label{eq. system dynamics 1}
\end{equation}
where $\mv q\in \mathbb{R}^n$ and $\mv \tau\in\R^n$ denote the generalized coordinate and the control input, respectively; $\mm M(\mv q)\in\mathbb{R}^{n\times n}$, $\mm C(\mv q,\dot {\mv q})\in\mathbb{R}^{n\times n}$, and $\mm G(\mv q)\in\mathbb{R}^{n}$ denote the unknown system function matrices; and $\mm F(\mv q)\in\mathbb{R}^{n}$ denotes the external force.
Using the user-designed matrices $\mm M_0>0,\mm C_0$ and $\mm G_0$, \eqref{eq. system dynamics 1} can be represented as 
\begin{equation}
    \mm M_0\ddot {\mv q}+\mm C_0\dot {\mv q}+\mm G_0 = \mv \tau + \mv f(\mv q,\dot {\mv q},\ddot {\mv q})
    \label{eq. system dynamics 2}
\end{equation}
where $\mv f(\mv q,\dot {\mv q},\ddot {\mv q}) \triangleq -(\mm M-\mm M_0)\ddot {\mv q}-(\mm C-\mm C_0)(\mv q,\dot {\mv q})\dot {\mv q} -(\mm G-\mm G_0) -\mm F(\mv q)$ denotes the residual unknown term.

Hence, the objective of the control design is to make $\mv q$ track the continuously differentiable desired trajectory $\mv q_d(t):\R\to \R^n$ under the unknown terms $\mv f$.

@@

The backstepping control-based approach is utilized to generate a reference signal $\mv z^*\triangleq -{k_q}\tilde {\mv q}+\dot {\mv q}_d$ for $\mv {z}\triangleq \dot {\mv q}$, where $\tilde {\mv q}\triangleq \mv {q} - {\mv q}_d$ and $k_q\in\R_{>0}$.
The desired stabilizing controller can be designed as follows:
\begin{equation}
    \mv \tau^* = 
    -\mm M_0\cdot ({k_z}\tilde {\mv z})
    + 
        ( 
            -\mm M_0\tilde {\mv q}+\mm C_0{\mv z}+\mm G_0-\mv f+\mm M_0 \dot {\mv z}^*
        )
    \label{eq. desired control}
\end{equation}
where $\tilde {\mv z}\triangleq \mv z-\mv z^*$ and $k_z\in\R_{>0}$.
Note that the desired controller cannot be realized because of $\mv f$.

To approximate the desired controller, an NN is utilized.
The NN with a single hidden layer is represented as 
\begin{equation}
    \mv \Phi({\mv q_{NN}};\mv \theta)\triangleq \mm V_1^\top \mv \phi(\mm V_0^\top {\mv q_{NN}})
\end{equation}
where ${\mv q_{NN}}\in\R^{l_0+1}$ denotes the NN input vector, $\mm V_i\in\R^{(l_i+1)\times l_{i+1}},\ i\in[0,1]$ denotes the weight matrix of the $i\textsuperscript{th}$ layer and $\mv \phi:\R^{l_1}\to\R^{l_1+1}$ denotes the activation function layer.
The activation function layer consists of an elementwise nonlinear function $\sigma(\cdot)$ and an augmented $1$ to combine the bias terms in the weight matrix (\ie $\mv \phi(\mv x) = [\sigma(\mv x_{(1)}),\cdots, \sigma(\mv x_{(l_1)}), 1]^\top $).
For further simplicity, let $\mv \theta\triangleq[\mv \theta_1^\top ,\mv \theta_0^\top ]^\top \in\R^{\Xi}$ denote the total weight vector, where $\mv \theta_i\triangleq \text{vec}(\mm V_i)\in\R^{\Xi_i}$ denotes the vectorized weights.
$\Xi_i=(l_i+1)\cdot l_{i+1}$ and $\Xi=\Xi_0+\Xi_1$ denote the number of each layer and total weights, respectively.

Using this NN, the desired controller $\mv \tau^*$ can be approximated by the ideal weight vector $\mv \theta^*$ for a compact subset $\Omega_{NN}\in\R^{l_0+1}$ to $\epsilon$ accuracy such that $\sup_{{\mv q_{NN}}\in\Omega_{NN}} \allowbreak \Vert\mv \Phi({\mv q_{NN}};\mv \theta^*) - \mv \tau^*\Vert=\epsilon<\infty$ \cite{RN35}.
% The boundedness of the norm of $\theta^*$  and approximation error $\epsilon$ are also claimed in \cite{RN35}.
The ideal weight vector $\mv \theta^*$ is typically assumed to be bounded.
Then, using the estimated weight vector $\hat{\mv \theta}=[\hat{\mv \theta}_1^\top ,\hat{\mv \theta}_0^\top ]^\top $ of $\mv \theta^*=[\mv \theta_1^{*\top},\mv \theta_0^{*\top}]^\top $, the desired controller $\mv \tau^*\approx -\mv \Phi(\mv q_{NN};\mv \theta^*)-\epsilon$ can be approximated as follows:
\begin{equation}
    \mv \tau \triangleq -\mv \Phi({\mv q_{NN}};\hat{\mv \theta}).
    \label{eq. control law}
\end{equation}
For further sections, let $\mv \Phi^*\triangleq\mv \Phi({\mv q_{NN}};\mv \theta^*)$ and $\mv \phi^* \triangleq\mv \phi(\mm V_0^{*\top}{\mv q_{NN}})$, and let $\hat{\mv \Phi}\triangleq{\mv \Phi}({\mv q_{NN}};\hat{\mv \theta})$, $\hat{\mv \phi} \triangleq\mv \phi(\hat {\mm V}_0^{\top}{\mv q_{NN}})$ and $\hat{\mv \phi}' \triangleq \partial \hat{\mv \phi}/\partial (\hat {\mm V}_0^\top \mv q_{NN})$.

Using \eqref{eq. system dynamics 2}, \eqref{eq. desired control}, and \eqref{eq. control law}, the error dynamics can be obtained as the first-order system of augmented error $\mv \xi\triangleq[\tilde {\mv q}^\top ,\tilde {\mv z}^\top ]^\top \in\R^{2n}$ as follows:
\begin{equation}
    \dot{\mv \xi} 
    = 
    \mm A_\xi{\mv \xi}+\mm B_\xi(\mv \Phi^*-\hat{\mv \Phi}+\epsilon)
\end{equation}
where 
\begin{equation}
    \mm A_\xi \triangleq 
    \begin{bmatrix}
        -{k_q} \mm I_{n\times n} &\mm I_{n\times n}
        \\
        -\mm I_{n\times n}& -{k_z} \mm I_{n\times n}
    \end{bmatrix}
    ,\ 
    \mm B_\xi \triangleq 
    \begin{bmatrix}
        \mm 0_{n\times n}\\\mm M_0^{-1}
    \end{bmatrix}.
\end{equation}

@@

Using the augmented error $\mv \xi$, the quadratic objective function is defined as
\begin{equation}
    J(\mv \xi;\hat{\mv \theta})\triangleq
    {1\over 2}
    \mv \xi^\top \mm W\mv \xi
\end{equation}
where $\mm W=\mm W^\top >0$ denotes the weighting matrix.

For the boundedness of the weights, weight norm constraints are imposed on the adaptation process such that $c_{\theta_i}=\Vert{\hat{\mv\theta}}_i\Vert^2-\bar\theta^2_i,\ i\in[0,1]$.
% Then the objective of the adaptation process can be reformulated into the constrained optimization problem as follows:
The control problem can be reformulated into a constrained optimization problem as follows:
\begin{equation}
    \begin{matrix}
        \min_{{\hat{\mv\theta}}} \ J(\mv \xi;{\hat{\mv\theta}})
        \\ \\
        \begin{aligned}
        \text{subject to\;}&c_{j}\le0, \quad j\in\mathcal I = [\theta_0,\theta_1].
        \end{aligned}
    \end{matrix}
\end{equation}
In this optimization problem, $\xi$ is considered a predefined parameter.
The corresponding Lagrangian function is defined as
\begin{equation}
    L(
        \mv \xi,
        {\hat{\mv\theta}},
        [\lambda_j]_{j\in\mathcal I}
    )
    \triangleq 
    J(\mv \xi;{\hat{\mv\theta}})
    +
    \sum_{j\in\mathcal I} \lambda_jc_j({\hat{\mv\theta}})
    .
\end{equation}
% where $\mathcal I=\{j\ \vert \ c_j\ge 0\}$ denotes the active set.
To solve the dual problem $\min_{{\hat{\mv\theta}}} \max_{[\lambda_j]_{j\in\mathcal I}}  L$, the adaptation law is derived as follows:
\begin{align}
        \dot {{\hat{\mv\theta}}}&=-\alpha {\partial L\over\partial {\hat{\mv\theta}}}
        =-\alpha 
        \bigg(
        {\partial J\over \partial {\hat{\mv\theta}}}+\sum_{j\in\mathcal{A}}
        \lambda_j {\partial c_j\over \partial {\hat{\mv\theta}}}
        \bigg),
    \label{eq. adaptation law th}
        \\
        \dot\lambda_j& = \beta_j{\partial L\over\partial \lambda_j} = \beta_j c_j ,
        \quad\quad\quad\quad      \      
        \forall j\in\mathcal I,
    \label{eq. adaptation law L}
        \\
        \lambda_j & = \max(\lambda_j,0) ,
        \quad\quad\quad\quad\ \ \ \ \ 
        \forall j\in\mathcal I,
    \label{eq. adaptation law L max}
\end{align}
where $\alpha\in\R_{>0}$ and $\beta_j\in\R_{>0}$ denote the adaptation gain and the update rate for each Lagrange multiplier, respectively.
\color{red}
% The Eq.~\eqref{eq. adaptation law L max} is analog to using an active set $\mathcal I=\{j\ \vert \ c_j\ge 0\}$ in constrained optimization theory which excludes terms of satisfied constraints in adaptation law.
\color{black}

Using the chain rule, the gradient of the objective function with respect to the weights (\ie $\partial J/\partial {\hat{\mv\theta}}$) in \eqref{eq. adaptation law th} can be represented as
\begin{equation}
    {\partial J\over\partial {\hat{\mv\theta}}}
    =
    {\partial \mv\xi\over\partial {\hat{\mv\theta}}}^\top 
    \mm W
    {\partial \mv\xi\over\partial {\hat{\mv\theta}}}
    .
\end{equation}
The calculation of $\partial J/\partial {\hat{\mv\theta}}$ is not straightforward because of the dynamics of $\xi$.
Using the forward sensitivity method presented in \cite{RN34}, the sensitivity equation can be obtained as follows:
\begin{equation}
    \dot {\mm \eta} =
    \mm A_\xi - \mm B_\xi 
    % (\partial \hat{\mv\Phi}/\partial {\hat{\mv\theta}})
    {\partial \hat{\mv\Phi}\over\partial {\hat{\mv\theta}}}
    ,
    \label{eq. sen eq}
\end{equation}
where ${\mm \eta}\triangleq \partial \mv\xi/\partial {\hat{\mv\theta}}\in\R^{2n\times \Xi}$ denotes the sensitivity of the weights to the augmented error.
The initial value of ${\mm \eta}$ is zero since the initial $\mv\xi$ is independent of the weights.
By decomposing for each layer, the dynamics of ${\mm \eta}_i\triangleq \partial \mv\xi/\partial {\hat{\mv\theta}}_i\in\R^{2n\times \Xi_i}$ can be represented as
\begin{equation}
    \begin{aligned}
        \dot{\mm \eta} =&
        \begin{bmatrix}
            {\mm \eta}_1&{\mm \eta}_0
        \end{bmatrix}'
        \\
        =&
        \mm A_\xi
        % \begin{bmatrix}
        [
            {\mm \eta}_1
            \ \ 
            {\mm \eta}_0
        ]
        % \end{bmatrix}
        -\mm B_\xi
        % \begin{bmatrix}
        [
            (\mm I_{l_2}\otimes \hat{\mv\phi}^\top )
            \ \ 
            \hat {\mm V}_1^\top \hat{\mv\phi}'(\mm I_{l_1}\otimes {\mv q_{NN}}^\top )
        ]
        % \end{bmatrix}
        .
    \end{aligned}
\end{equation}
The calculation of $\partial \hat{\mv\Phi}/\partial{\hat{\mv\theta}}$ is introduced in \cite{RN36}.
In conclusion, the gradient of the objective with respect to the weights can be obtained as ${\partial J/\partial {\hat{\mv\theta}}} = {\mm \eta}^\top\mm W\mv\xi$ by simulating the sensitivity equation \eqref{eq. sen eq}.

On the other hand, the gradient of the constraints with respect to the weights can be represented as follows:
\begin{equation}
    \frac{\partial c_{\theta_0}}{\partial{\hat{\mv\theta}}}
    = 
    \begin{bmatrix}
        \mm 0_{\Xi_1\times 1} \\
        2{\hat{\mv\theta}}_0 
    \end{bmatrix}
    ,\quad
    \frac{\partial c_{\theta_1}}{\partial{\hat{\mv\theta}}}    
    = 
    \begin{bmatrix}
        2{\hat{\mv\theta}}_1 \\
        \mm 0_{\Xi_0\times 1}
    \end{bmatrix}
    .
\end{equation}

@@

\begin{theorem}
For the dynamical system in \eqref{eq. system dynamics 1}, the proposed controller \eqref{eq. control law} and the adaptation law \eqref{eq. adaptation law th}, \eqref{eq. adaptation law L} and \eqref{eq. adaptation law L max} ensure the boundedness of the tracking error $\mv \xi$ and the weight estimation ${\hat{\mv\theta}}$, provided that the control gains ${k_q}$ and ${k_z}$ satisfy \eqref{eq. ctrl stable condition}.
\end{theorem}

@@

\subsection*{Step 1: Boundedness of ${\hat{\mv\theta}}_1,{\mm \eta}_1,\mv \xi$}

@@

The dynamics of $\mv \xi$ can be represented as
\begin{equation}    
    \dot {\mv\xi} = \mm A_\xi\mv \xi +\mm B_\xi
    (
        -\hat {\mm V}_1^\top \hat{\mv\phi}+\mv w(t)
    )
\end{equation}
where $\mv w(t)\triangleq {\mm V}_1^{*T}\mv\phi^*+\mv \epsilon$ is a lumped residual term, which is bounded as $\Vert\mv  w(t)\Vert\le \bar w< 0$ since $\Vert\mv\theta_1^*\Vert$ and $\Vert\mv\phi^*\Vert$ are bounded.
The dynamics of ${\mm \eta}_1$ and ${\hat{\mv\theta}}_1$ are represented as
\begin{equation}
    \begin{aligned}
        \dot{\mm \eta}_1 =
        & 
        \mm A_\xi {\mm \eta}_1 -\mm B_\xi (\mm I_{l_{2}}\otimes \hat{\mv\phi}^\top )
        \\
        \dot{{\hat{\mv\theta}}}_1 =
        & -\alpha 
        (
            {\mm \eta}_1^\top \mm W\mv\xi+2\lambda_{\theta_1} {\hat{\mv\theta}}_1
        ).
    \end{aligned} 
\end{equation}
According to \cite[Chap.~4 T.~1.9]{RN32}, the boundedness of ${\mm \eta}_1$ can be obtained since $\mm A_\xi$ is stable and the residual term $-\mm B_\xi (\mm I_{l_{2}}\otimes \hat{\mv\phi}^\top )$ is bounded.
